A Temporal Logic Based Framework for Intrusion Detection

### 2023-09-08 Paper/main.tex
update

%% INTRODURRE VELOCE ESEMPIO E STRUTTURA DEL PROGRAMMA

%%%%%   DA RISCRIVERE ALLA FINE
In Section 2 we will analyze the basics of EAGLE. In Section 3 we will analyze MONID framework and some evalutations on the performances. In Section 4 we will see some attacks specifications. In section 5 we will conclude.
%%%%5

@@

It's essential to emphasize that the logic EAGLE operates with a \textit{finite} trace semantics; the monitoring algorithm, as written before, can check whetet the formula is satisfied or not only at the conclusion of a trace. However, in intrusion detection scenarios, the notion of the "end of a trace" makes no sense since event sequences can theoretically be infinite. In such cases, the aim is to trigger an alarm as soon as a property is violated and this is done by continuously checking the formula's satisfaction status after each event. In practice, once we have the specification of the attack pattern in terms of $\varphi$, we can express the system's safe behavior as $\square(\neg \varphi)$. Note that if a sequence of events corresponding to the attack is discovered, this formula becomes false.

@@

\section{Conclusions}

>>>>

%% INTRODURRE VELOCE ESEMPIO E STRUTTURA DEL PROGRAMMA
%%%%%   DA RISCRIVERE ALLA FINE
In Section 2, we'll deeply explore EAGLE, covering everything from its syntax to semantics and the evaluation algorithm, with the add of practical examples. Moving on to Section 3, we'll delve into the MONID framework, which employs EAGLE to construct an intrusion detection system. Section 4 will focus on specific EAGLE rules tailored for the detection of various types of attacks and finally, in Section 5, we will draw our conclusions.
%%%%5

@@

It's essential to emphasize that the logic EAGLE operates with a \textit{finite} trace semantics; the monitoring algorithm, as written before, can check whether the formula is satisfied or not only at the conclusion of a trace. However, in intrusion detection scenarios, the notion of the "end of a trace" makes no sense since event sequences can theoretically be infinite. In such cases, the aim is to trigger an alarm as soon as a property is violated and this is done by continuously checking the formula's satisfaction status after each event. In practice, once we have the specification of the attack pattern in terms of $\varphi$, we can express the system's safe behavior as $\square(\neg \varphi)$. Note that if a sequence of events corresponding to the attack is discovered, this formula becomes false.

@@

\section{Conclusions}
In conclusion, we have introduced a new temporal logic named EAGLE, which serves as the foundation for the MONID framework, which is a useful tool for system intrusion detection.
MONID continuously monitor these attack specifications during system execution and when the observed execution contradicts the monitored formula, we detect an intrusion. We have showcased this by creating formulas to identify several common attacks and testing our monitoring algorithm on large event logs provided by DARPA. These examples can serve as templates for specifying many other types of attacks. The obtained results were excellent, both in terms of detected attacks and system overhead.
